---
title: "Addressing Subposterior Conflict With Weighted Proposals"
author: "Andrew Manderson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
fontfamily: tgpagella
fontsize: 10pt
papersize: a4
geometry: margin=2.25cm
bibliography: ../0bibliography/year-1-bib.bib
csl: aam71-test.csl
output: 
  pdf_document:
    includes:
      in_header:
        tex-input/pre.tex
    fig_caption: true
    number_sections: true
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = NA, out.width = "100%", fig.align = "center", auto_pdf = TRUE, cache = TRUE, fig.width = 5, fig.asp = 0.5)
sim_pars <- readRDS("rds/norm-norm-ex/sim-pars.rds")
attach(sim_pars) # Needed to dynamically generate table
```



The need for accurate self-density ratio estimates arose from a conflict between $m^{\text{th}}$ stage posterior and the unknown prior marginal distribution, particularly when these things disagreed over the scale of the $(m + 1)^{\text{th}}$ stage posterior.
However, this was just a disagreement of scale, the distributions in question were not disjoint.
Here, we show the issue of disjoint subposteriors and suggest / demonstrate a potential solution in for Markov melding.

# Manifestations of conflict: disjoint posteriors

Consider the two subposteriors $\pd_{1}(\phi \mid Y_{1})$, $\pd_{2}(\phi \mid Y_{2})$, and melded posterior $\pmeld(\phi \mid Y_{1}, Y_{2})$ shown in Figure&nbsp;\ref{fig:subpost_disagreement}.
Using the samples of the subposteriors as a proposal for the overall posterior results in a degenerate sample of the overall posterior, due to an insufficient number of samples in the support of $\pmeld(\phi \mid, Y_{1}, Y_{2})$.
Sampling a weighted version of one of the subposteriors $\pd_{1}(\phi \mid Y_{1})\text{u}_{1}(\phi)$, the green line in Figure&nbsp;\ref{fig:subpost_disagreement}, is the idea explored here. 

```{r subpost_disagreement, fig.cap = "Conflicting subposteriors."}
knitr::include_graphics("plots/intro/subpost-disagreement.pdf")
```

# Methods - Maths

Consider the overall melded target posterior for $\Nm = 2$ models
\input{tex-input/method/0010-melded-target-posterior.tex}
The stage one target is then
\input{tex-input/method/0020-stage-one-target.tex}
Hence, the stage two acceptance probability is 
\input{tex-input/method/0030-stage-two-acceptance.tex}

The presence of the $\tarw_{1}(\phi)$ terms in the stage 2 acceptance probability implies that we have some degree of control over the acceptance rate.
The $\tarw_{1}(\phi) \pd_{1}(\phi)$ term also suggests that this idea may be more linked than we suspect to our weighted estimation methodology for the prior marginal distribution.

For $\Nm > 2$, we will accrue $\tarw_{\modelindex}(\phi)$ terms, which may be problematic.

## How to choose $\text{u}(x)$

- is $\text{u}(x)$ the right symbol?
    - the alphabetical successor was $\text{t}(x)$, but this looks like a t-distribution
- What is the method here?
    - Look at subposteriors
    - Look at stage one samples
    - do enough of the latter lie in the distributional product of the former? Can we formalise this?
- What mathematical tools do we need to choose a $\text{u}(x)$ that isn't itself a distribution, what properties do we require?
    - The idea here would be to choose something that widens the target (that isn't tempering)
    - Really we want to map the target distribution to the maximum entropy distribution (uniform), which involves knowing the inverse mapping
    - choose $\text{u}(x) = 1 \mathop{/} f(x; \mu, \sigma^2)^{\Nx - k}$, where k is some safety factor, and $f(x; \mu, \sigma^2)$ is the normal density.
        - Choose $\mu = \overline{y}$?
        - This deals with the narrowing, but doesn't address the translation
        - $\text{u}(x) = f(x; \overline{y}_{2}, \hat{s}_{2}^2) \mathop{/} f(x; \overline{y}_{1}, \hat{s}_{1}^2)^{\Nx - k}$ can address both the narrowing
        and the translation (sort of).
            - what are the conditions for this to be sensible (heavy-ness of tails).
            - This is nice because it only needs summary statistics, there is no point building a hugely complicated function that needs to look at all of the data?
    - ~~Might as well choose $\text{u}(x) = x^{-(\Nx - k)}$?~~
    - This is going to get overwhelmed with numerical error fairly quickly.
    - Feels a lot like importance sampling, can we use the importance sampling results to justify approaches like this?
        - PSIS paper has some good ideas about when IS works / fails
    - What about the control variates idea used in subsampling MCMC, they use them to stabilise the variance of the log-posterior, but we could use them to inflate it? (I don't know about that).

## Potential issues

- We might get bitten by distributional concentration again (introducing another multiplicative term narrows the target), have to see
- Feasible in 1/2/(3?) dimensions, and melding cases where $\Nm$ is $<10$?. 
    - Trying to think of cases where it is easyish to figure out where the overall posterior will be by visual inspection.
    - More general idea for finding "midpoint" between samples from two distributions.
        - Seems a lot like the idea of Barycentres in: https://arxiv.org/abs/1508.05880
- Can use ESS as a diagnostic for particle degeneracy 
 
# Example - norm norm

Starting from the same Normal-Normal model used to demonstrate the importance of accurately estimating the prior marginal distribution, we adjust some of the hyperparmeters to manufacture disjoint subposteriors.
We have $\Nm = 2$ models, with common parameter $\phi$,
\input{tex-input/norm-norm-ex/0010-norm-models.tex}
Choosing the following values for the hyperparmeters and number of data points

| Parameter             | Value           |
|-----------------------|-----------------|
| $\Nx_{1}$             | `r n_y_1`       |
| $\Nx_{2}$             | `r n_y_2`       |
| $\varepsilon_{1}^{2}$ | `r sigma_1`     |
| $\varepsilon_{2}^{2}$ | `r sigma_2`     |
| $\mu_{1}$             | `r mu_phi_1`    |
| $\mu_{2}$             | `r mu_phi_2`    |
| $\sigma_{1}^2$        | `r sigma_phi_1` |
| $\sigma_{2}^2$        | `r sigma_phi_1` |

results in the subposterior distributions in Figure&nbsp;\ref{fig:subposteriors}.

```{r subposteriors, fig.cap = "subposteriors and target"}
knitr::include_graphics("plots/norm-norm-ex/subposteriors.pdf")
```

## Stage one target and $\text{u}_{1}(\phi)$.

- Choice is slightly weird
- need to define some notation for the augmented target?

```{r u_func_augmented_target, fig.cap ="$\\text{u}_{1}(\\phi)$ for various choices of $k$, and the resulting augmented target (log scale, unnormalised)."}
knitr::include_graphics("plots/norm-norm-ex/u-function-augmented-target.pdf")
```


## Numerical tests

We want to compare the melded posterior distribution sampled using the following methods:

####  sequential MH, with no $\text{u}_{1}(\phi)$ (normal melding)

Figure&nbsp;\ref{fig:no_u_traces}.

```{r no_u_traces, fig.cap = "$n$-stage traceplots, no subposterior weighting"}
knitr::include_graphics("plots/norm-norm-ex/no-u/stage-traces.pdf")
```


####  sequential MH, _with_ $\text{u}_{1}(\phi)$ (Idea developed here)

Figure&nbsp;\ref{fig:with_u_traces}.
Using $\text{u}(x) = f(x; \overline{y}_{2}, \hat{s}_{2}^2) \mathop{/} f(x; \overline{y}_{1}, \hat{s}_{1}^2)^{\Nx - 3}$ for the moment.

Not perfect, still thinking. Also don't think I've quite got the implementation & maths correct just yet.
Need to think more about the _right_ weighting function, but current setup allows for easy experimentation.

```{r with_u_traces, fig.cap = "$n$-stage traceplots, \\textit{with} subposterior weighting."}
knitr::include_graphics("plots/norm-norm-ex/with-u/stage-traces.pdf")
```

####  joint MH, i.e. sample $\pmeld(\phi \mid Y_{1}, Y_{2})$ all at once (it is possible here, and I'm interested in the results we get)

Figure&nbsp;\ref{fig:joint_trace}

```{r joint_trace, fig.cap = "trace of the melded posterior, sampled by MCMC directly."}
knitr::include_graphics("plots/norm-norm-ex/joint/trace.pdf")
```

Comparing to Figure&nbsp;\ref{fig:with_u_traces} allows us to see we don't have the weighting function quite right yet, we have gone to far towards the second subposterior.

####  Analytical truth (I _think_ this is still easy enough to calculate with $\Nx = 5$).

TBA - need to do maths & implement this calculation for arbitrary number normal distributions / means.

_Sun 7 Jul 15:47:26 2019_: I was very wrong about this. The algebra is slightly fiddly, but the numerics of this are just terrible. Have not yet figured out where I'm introducing the error.

# Example - real world? Where can we get disjoint conflict from.

- Investigate the Carlin example Rob sent

<!-- -------------------- END OF MAIN BODY OF DOCUMENT -------------------- -->
\newpage

<!-- The {-} tag here suppresses the section numbering. -->
# Bibliography {-}

<!-- This makes pandoc-citeproc put the references before the end of document. -->
<div id="refs"></div>

\newpage

<!-- Now switch to alphabetical numbering for the appendix, and reset the counter. -->
\renewcommand{\thesection}{\Alph{section}}
\setcounter{section}{0}

# Appendix 